---
title: "Apprentissage Statistique"
subtitle: "Application"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(knitr)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, eval=FALSE)
```

```{r Libraries, include=FALSE}
library(dplyr)
library(stats)
library(tidyr)
library(lubridate)
library(stringr)
library(gridExtra)
library(readr)
library(e1071)
library(kernlab)
library(lattice)
library(tidyr)
library(ggplot2)
library(ggmap)
library(psych)
library(GGally)
library(lubridate)
library(stringi)
library(forcats)
library(stringr)
library(corrplot)
```

We will work with the dataset of Airbnb for the city of Paris

```{r data importation}
data <- read.csv("https://raw.githubusercontent.com/lavergnetse/Data/master/listings_Paris.csv", sep = ",", stringsAsFactors = TRUE)

glimpse(data)
```

## Data pre-processing

### 1) Keep only the observations located inside of Paris. You can use the variable *zipcode* which indicate the department to help you. 

```{r question_1, exercise=TRUE}
# Overview of the zipcode

# Extraction of the department number

# Eliminate observations that are not in the department 75

```

```{r question_1-solution}
# Overview of the zipcode
summary(data$zipcode)

# Extraction of the department number
data$department <- str_sub(as.character(data$zipcode), start = 1, end = 2)

# Eliminate observations that are not in the department 75
data <- data[data$department=="75",]
```


### 2) Gather in a new vector the variables that give informations about the housing (price, location, attributes, rating), the host attributes and the rental rules. Keep the *zipcode* as an indicator of the district (last two numbers). Compile a new dataframe which contain only these variables.   

```{r question_2, exercise=TRUE}

```

```{r question_2-solution}
listings_keep <- c("id", "price", "host_is_superhost", "zipcode",
                   "latitude","longitude", "property_type", "accommodates",
                   "bedrooms", "beds", "cleaning_fee",
                   "minimum_nights", "availability_365",
                   "review_scores_rating","cancellation_policy")

data <- data[,listings_keep]
```


### 3) Create a new variable called "arrondissement" that give informations about the district. Keep only the observations that belong to arronssements 1 to 20.

```{r question_3, exercise=TRUE}
# Create the new variable

# Overview of the variable

# Keep only the observation with valid arrondissement

# Remove the column zipcode 


```

```{r question_3-solution}
# Create the new variable
data$arrondissement <- str_sub(as.character(data$zipcode), start = 4, end = 5)

# Overview of the variable
summary(as.factor(data$arrondissement))

# Keep only the observation with valid arrondissement
districts <- c("", "00")
data <- data[!data$arrondissement %in% districts,]
data$arrondissement[data$arrondissement == "8"] <- "08"

# Remove the column zipcode 
data <- data[,-which(names(data) == "zipcode")]
```


### 4) Is there any duplicated observations ? Eliminate them if they exist.

```{r question_4, exercise=TRUE}

```

```{r question_4-solution}
sum(duplicated(data$id))
```


### 5) Remove the $ sign for both *price* and *cleaning_fee*.

```{r question_5, exercise=TRUE}

```

```{r question_5-solution}
data$price = as.numeric(gsub("[\\$,]", "", data$price))

data$cleaning_fee = as.numeric(gsub("[\\$,]", "", data$cleaning_fee))
```


### 6) Remove the observations where the price is null. Replace by NA's the missing values that are represented by empty cells.

```{r question_6, exercise=TRUE}
# Price

# See which columns have empty cells

# Replace the empty cases by NA's (useful function that automatically convert the factor
# into character)

```

```{r question_6-solution}
# Price
data <- data %>% filter(price > 0) 

# See which columns have empty cells
empty_cells <- as.data.frame(colSums(data == ""))
empty_cells

# Replace the empty cases by NA's (useful function that automatically convert the factor
# into character)
data <- data %>% mutate_at(.vars = c("host_is_superhost", "cancellation_policy"), 
                           .funs = list(~ifelse(.=="", NA, as.character(.))))

```


### 7) In wich other variables is there NA ?  

```{r question_7, exercise=TRUE}
# NA count

```

```{r question_7-solution}
# NA count 
na_count <- as.data.frame(colSums(data%>%is.na(.)))
na_count
```


### 8) Replace the NA in the discrete variables by performing median imputation. For the categrorial ones, replace NA simply by the mode (the most frequent value). Finally for *cleaning_fee*, set to 0 the observation with NA. 

```{r question_8, exercise=TRUE}
#  === Bedrooms === 

#  === Beds === 

# === Host is superhost ===

# === Review scores rating ===

# === Cleaning fee ===


```

```{r question_8-solution}
#  === Bedrooms === 

data <- data %>%
  group_by(
    property_type,
    accommodates) %>%
  mutate(bedrooms=ifelse(is.na(bedrooms),
                         median(bedrooms,na.rm = T),bedrooms)) %>%
  ungroup()

#  === Beds === 

data <- data %>%
  group_by(
    property_type,
    accommodates,
    bedrooms) %>%
  mutate(beds=ifelse(is.na(beds),
                         median(beds,na.rm=T),beds)) %>%
  ungroup()

# === Host is superhost ===

data[is.na(data$host_is_superhost),"host_is_superhost"] <- "f"

# === Review scores rating ===

data <- data %>%
  group_by(arrondissement,
           accommodates,
           property_type,
           host_is_superhost) %>%
  mutate(review_scores_rating=ifelse(is.na(review_scores_rating), 
                                     median(review_scores_rating,na.rm=T),
                                     review_scores_rating)) %>%
  ungroup()


# === Cleaning fee ===

data[is.na(data$cleaning_fee), "cleaning_fee"] <- 0
```


### 9) Are there still NAs ? If yes remove these observations.

```{r question_9, exercise=TRUE}

```

```{r question_9-solution}
sum(is.na(data))
data <- na.omit(data)
```


### 10) Create subgroup for the categorical variables : 

- 4 subgroups for *property_type*: Entire_home ; Private_room ; Shared_room and other
- 2 subgroups for *cancellation_policy* : strict vs flexible

```{r question_10, exercise=TRUE}
# # ===== ===== Property type ===== =====

# We groups all 32 subgroups of housing types into 3 big groups

# # ===== ===== Cancellation policy ===== =====

```

```{r question_10-solution}
# # ===== ===== Property type ===== =====

# We groups all 32 subgroups of housing types into 3 big groups

Entire_home <- c("Aparthotel","Apartment", "Condominium", "Earth house", 
                 "Serviced apartment", "Loft", "Dome house","Tiny house", 
                 "Townhouse", "House", "Villa")

Shared_room <- c("Dorm", "Hostel", "Guesthouse")

Private_room <- c("Bed and breakfast", "Boutique hotel", "Hotel",
                  "Guest suite", "In-law")

Other <- c("Barn","Boat", "Bungalow", "Cabin", "Campsite", "Cave", "Cottage", 
           "Farm stay", "Houseboat", "Island", "Nature lodge", "Other", "Plane")

data$property_type <- as.character(data$property_type)

data <-
  mutate(data,
         property_type = ifelse(property_type %in% Entire_home,
                                "Entire_home", property_type),
         property_type = ifelse(property_type %in% Shared_room,
                                "SharedRoom", property_type),
         property_type = ifelse(property_type %in% Private_room,
                                "PrivateRoom", property_type),
         property_type = ifelse(property_type %in% Other,
                                "Others", property_type))


# # ===== ===== Cancellation policy ===== =====

strict <- c("strict_14_with_grace_period", "super_strict_30", "super_strict_60", 
            "strict", "luxury_super_strict_95")

flexible <- c("flexible", "moderate", "luxury_moderate")

data$cancellation_policy <- as.character(data$cancellation_policy)

data <-
  mutate(data,
         cancellation_policy = ifelse(cancellation_policy %in% strict,
                                "Strict", cancellation_policy),
         cancellation_policy = ifelse(cancellation_policy %in% flexible,
                                "Flexible", cancellation_policy))
```


### 11) Download the new dataframe

```{r question_11, exercise=TRUE}

```

```{r question_11-solution, eval=FALSE}
write.csv(data, "/Users/.../paris_data.csv", row.names = FALSE)
```


## Data Analysis

```{r import cleaned data, eval=FALSE}
data <- read.csv("https://raw.githubusercontent.com/lavergnetse/Data/master/paris_data.csv", sep = ",", stringsAsFactors = TRUE)
``` 

### 1) Look at the distribution of the property type and the location (*arrondissement*).

```{r question_12, exercise=TRUE}

```

```{r question_12-solution}
data$arrondissement <- as.factor(data$arrondissement)
interest_varables <- c("arrondissement", "property_type")

library(ggthemes)

for (i in interest_varables) {
  plot <- ggplot(mapping = aes_string(x = fct_infreq(data[,i]),
                                     fill = data[,i])) +
    geom_bar(width = 1, colour = "black", show.legend = F,
             aes(y = ((..count..)/sum(..count..)))) +
    scale_y_continuous(labels = scales::percent) +
    theme_economist() + labs(x = "", y = "") +
    ggtitle(str_replace_all(str_to_title(i), c("_" = " ")))
  print(plot)
}
```


### 2) Perform the logarithmic transformation of the price. Then compare the price by property type.

```{r question_13, exercise=TRUE}

```

```{r question 13-solution}
data$price <- log(data$price) 

ggplot(data, aes(x = property_type, y = price)) +
  geom_boxplot(outlier.colour = "darkblue", outlier.size = 0.5,
               color="deepskyblue", fill="cyan", alpha=0.2) +
  ggtitle("Average Airbnb prices per night by property type") +
  labs(x = "", y = "Average prices per night (in log)") + 
  theme_economist()
```


### 3) Remove the outliers for *price*. Compare one more time the price by property type.

```{r question_14, exercise=TRUE}
# Compute upper & lower limit

```

```{r question 14-solution}
Q <- quantile(data$price, probs=c(.25, .75))

# Compute upper & lower limit
upper_limit <- Q[2] + 1.5 * IQR(data$price)
lower_limit <- Q[1] - 1.5 * IQR(data$price)
data <- data %>% filter (price <= upper_limit & price >= lower_limit)

# New price
ggplot(data, aes(x = property_type, y = price)) +
  geom_boxplot(outlier.colour = "darkblue", outlier.size = 0.5,
               color="deepskyblue", fill="cyan", alpha=0.2) +
  ggtitle("Average Airbnb prices per night by property type") +
  labs(x = "", y = "Average prices per night (in log)") + 
  theme_economist()
```


### 4) Represent on a map the price per arrondissement.

```{r question_15, exercise=TRUE, message=FALSE}

```

```{r question 15-solution}
# Define the baseline (play with your parameters such as: min(longitude), max(latitude), 
# the level of zoom, ...)
borders <- c(bottom  = 48.8,
             top     = 48.9,
             left    = 2.25,
             right   = 2.4)

map <- get_stamenmap(borders, zoom = 14, maptype = "terrain", crop = FALSE)

secteur <- data %>% group_by(arrondissement) %>%
  summarize(longitude = median(longitude),
            latitude = median(latitude),
            prix_moyen = mean(price))

# Construct the map
ggmap(map) +
  geom_point(secteur, shape = 19,
             mapping = aes(x = longitude, y = latitude,
                           col = prix_moyen)) +
  scale_colour_gradient(low = "blue", high = "red") +
  theme(plot.title = element_text(hjust = 0.5),
               plot.background = element_rect(fill = "#BFD5E3")) +
  labs(x = "", y = "", col = "Average price") +
  ggtitle("Average price by Arrondissement") + 
  geom_label(secteur,
             mapping = aes(longitude, latitude,
                           label = arrondissement),
             size = 1.8, fontface = "bold",
             nudge_x = 0, nudge_y = 0.005)
```


### 5) Remove the variables that are not relevant for prediction. Then create dummies for *host_is_superhost*, *property_type* and *cancellation_policy*.

```{r question_16, exercise=TRUE}

```

```{r question_16-solution}
library(caret)

# Create a subset of the data to use in prediction
data_model <- data %>% select(-c("id", "arrondissement", "latitude", "longitude"))

# Dummies
dummies <- dummyVars(price ~., data_model, sep = "_")
new_set <- as.data.frame(predict(dummies, newdata = data_model))
data_model <- cbind(data_model$price, new_set)
```

### 6) Test/Train set

```{r question_17, exercise=TRUE}

```

```{r question_17-solution}
colnames(data_model)[which(names(data_model) == "data_model$price")] <- "price"

trainIndex <- createDataPartition(data_model$price, 
                                  p = .8, 
                                  list = FALSE, 
                                  times = 1)

data_train <- data_model[trainIndex,]
data_test  <- data_model[-trainIndex,]
```

